*Symbiodinium* communities in reef corals of St. John, U.S.V.I. (2012)

### 2016-08-17 analysis/STJ2012_analysis.Rmd
add 97% within-species OTU analysis

# Data pre-processing
### Import dataset
The [phyloseq](https://joey711.github.io/phyloseq/) package is used here to combine the OTU table, taxonomic assignments, and sample metadata into a single R data object (class 'phyloseq') to facilitate downstream analyses.
```{r data_import, cache=T}
# Import OTU tables
OTU97 <- otu_table(read.table("data/97_otus.tsv", header=T, check.names=F, row.names=1,
                              skip=1, sep="\t", comment.char=""), taxa_are_rows=T)
OTU97bs <- otu_table(read.table("data/97_otus_bysample.tsv", header=T, check.names=F, row.names=1,
                              skip=1, sep="\t", comment.char=""), taxa_are_rows=T)
OTU100 <- otu_table(read.table("data/100_otus.tsv", header=T, check.names=F, row.names=1,
                               skip=1, sep="\t", comment.char=""), taxa_are_rows=T)
# Import taxonomy
TAX97 <- tax_table(as.matrix(read.table("data/97_otus_tax_assignments.txt", sep="\t", row.names=2,
                                        col.names=c("Clade", "OTU.ID", "Taxonomy", "Eval", "Subtype"))))
TAX97bs <- tax_table(as.matrix(read.table("data/97_otus_bysample_tax_assignments.txt", sep="\t", row.names=2,
                                        col.names=c("Clade", "OTU.ID", "Taxonomy", "Eval", "Subtype"))))
TAX100 <- tax_table(as.matrix(read.table("data/100_otus_tax_assignments.txt", sep="\t", row.names=2,
                                         col.names=c("Clade", "OTU.ID", "Taxonomy", "Eval", "Subtype"))))
# Import sample data
SAM <- sample_data(read.delim("data/mapping_file.txt", sep="\t", header=T, row.names=1))
# Build phyloseq objects
phy97 <- phyloseq(OTU97, TAX97, SAM)
phy97bs <- phyloseq(OTU97bs, TAX97bs, SAM)
phy100 <- phyloseq(OTU100, TAX100, SAM)
```

### Filter dataset
```{r filter_dataset, cache=T}
# Filter OTUs by minimum count
# Set threshold count
n <- 10
# Identify OTUs below threshold count
taxa97 <- taxa_sums(phy97)[which(taxa_sums(phy97) >= n)]
taxa97bs <- taxa_sums(phy97bs)[which(taxa_sums(phy97bs) >= n)]  
taxa100 <- taxa_sums(phy100)[which(taxa_sums(phy100) >= n)]
# Remove taxa below threshold count
phy97.f <- prune_taxa(names(taxa97), phy97)
phy97bs.f <- prune_taxa(names(taxa97bs), phy97bs)
phy100.f <- prune_taxa(names(taxa100), phy100)

# Filter samples by minimum count
# Set threshold number of reads
sn <- 200
# Remove samples with fewer reads than threshold
phy97.f <- prune_samples(sample_sums(phy97.f)>=sn, phy97.f)
phy97bs.f <- prune_samples(sample_sums(phy97bs.f)>=sn, phy97bs.f)
phy100.f <- prune_samples(sample_sums(phy100.f)>=sn, phy100.f)
```
* Minimum count to retain OTU: `r n`
* Minimum count to retain sample: `r sn`

@@

#### Filtered dataset
```{r phy.f_histograms, cache=T, fig.height=4, fig.width=8}
# Compute summary statistics
stats97.f <- data.frame(`97% OTUs`=t(phystats(phy97.f)), check.names=F)
stats97bs.f <- data.frame(`97% within-sample OTUs`=t(phystats(phy97bs.f)), check.names=F)
stats100.f <- data.frame(`100% OTUs`=t(phystats(phy100.f)), check.names=F)

# Create and plot histograms
taxhist97 <- hist(log10(taxa_sums(phy97.f)), plot=F)
taxhist97bs <- hist(log10(taxa_sums(phy97bs.f)), plot=F)
taxhist100 <- hist(log10(taxa_sums(phy100.f)), plot=F)
samhist97 <- hist(log10(sample_sums(phy97.f)), plot=F)
samhist97bs <- hist(log10(sample_sums(phy97bs.f)), plot=F)
samhist100 <- hist(log10(sample_sums(phy100.f)), plot=F)

par(mfrow=c(2, 3), mar=c(3,3,1,1))
plot(taxhist97, col="black", main="97% OTU counts", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. OTUs", cex.lab=0.75, cex.axis=0.75)
plot(taxhist97bs, col="black", main="97% OTU counts", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. OTUs", cex.lab=0.75, cex.axis=0.75)
plot(taxhist100, col="black", main="100% OTU counts", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. OTUs", cex.lab=0.75, cex.axis=0.75)
plot(samhist97, col="black", main="97% OTU reads per sample", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. samples", cex.lab=0.75, cex.axis=0.75)
plot(samhist97bs, col="black", main="97% OTU reads per sample", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. samples", cex.lab=0.75, cex.axis=0.75)
plot(samhist100, col="black", main="100% OTU reads per sample", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. samples", cex.lab=0.75, cex.axis=0.75)

# Create stats table
knitr::kable(cbind(stats97.f, stats97bs.f, stats100.f))
```

#### Raw dataset
```{r phy_stats, cache=T, fig.height=4, fig.width=8}
# Compute summary statistics
stats97 <- data.frame(`97% OTUs`=t(phystats(phy97)), check.names=F)
stats97bs <- data.frame(`97% within-sample OTUs`=t(phystats(phy97bs)), check.names=F)
stats100 <- data.frame(`100% OTUs`=t(phystats(phy100)), check.names=F)

@@

@@

par(mfrow=c(2, 3), mar=c(3,3,1,1))
plot(taxhist97, col="black", main="97% OTU counts", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. OTUs", cex.lab=0.75, cex.axis=0.75)
plot(taxhist97bs, col="black", main="97% within-sample OTU counts", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. OTUs", cex.lab=0.75, cex.axis=0.75)
plot(taxhist100, col="black", main="100% OTU counts", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. OTUs", cex.lab=0.75, cex.axis=0.75)
plot(samhist97, col="black", main="97% OTU reads per sample", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. samples", cex.lab=0.75, cex.axis=0.75)
plot(samhist97bs, col="black", main="97% within-sample OTU reads per sample", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. samples", cex.lab=0.75, cex.axis=0.75)
plot(samhist100, col="black", main="100% OTU reads per sample", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. samples", cex.lab=0.75, cex.axis=0.75)

# Create stats table
knitr::kable(cbind(stats97, stats97bs, stats100))
```

### Transform count data
Count data are transformed to both relative abundance (proportions) and square-root proportions for downstream statistical analyses.
```{r transform}
# Convert to proportion (relative abundance)
phy97.f.p <- transform_sample_counts(phy97.f, function(x) x/sum(x))
phy97bs.f.p <- transform_sample_counts(phy97bs.f, function(x) x/sum(x))
phy100.f.p <- transform_sample_counts(phy100.f, function(x) x/sum(x))
# Apply transformation function
transform <- function(x) sqrt(x/sum(x))  # Set transformation function
phy97.f.t <- transform_sample_counts(phy97.f, transform)  # Transform data
phy97bs.f.t <- transform_sample_counts(phy97bs.f, transform) 
phy100.f.t <- transform_sample_counts(phy100.f, transform)
```

@@

## *Diploria strigosa*  
Notice how 97% OTUs and 97% within-sample OTUs are **NOT THE SAME!** Specifically, notice how in the 97% OTU dataset, all samples except two are dominated by the same OTU, denovo235. In the 97% within-sample OTU dataset, those samples are dominated by three different OTUs -- denovo73, denovo231, and denovo10. Look at the 100% OTU dataset to see the composition of unique sequence variants in these samples. Indeed, these samples are dominated by **different sequence variants**. However, in the 97% OTU approach, they have all been collapsed into the same OTU, while in the 97% within-sample clustering approach, samples with different dominant sequence variants are maintained as separate OTUs, even though these different sequence variants are more than 97% similar to each other. I suggest that the  within-sample clustering approach is a better way of treating *Symbiodinium* ITS2 sequence data.
```{r dstr, cache=T, fig.height=12, fig.width=8}
# Create subsetted phyloseq objects for Diploria strigosa
dstr97.f.p <- subset_samples(phy97.f.p, Species=="strigosa")
dstr97bs.f.p <- subset_samples(phy97bs.f.p, Species=="strigosa")
dstr100.f.p <- subset_samples(phy100.f.p, Species=="strigosa")
# Plot custom barplots for Diploria strigosa
par(mfrow=c(3,1), mar=c(4,4,2,0), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(dstr97.f.p, main="97% OTUs")
otubarplot(dstr97bs.f.p, main="97% within-sample OTUs")
otubarplot(dstr100.f.p, main="100% OTUs")
```

@@

## *Porites furcata*
```{r pfur, cache=T, fig.height=12, fig.width=8}
# Create subsetted phyloseq objects for Porites furcata
pfur97.f.p <- subset_samples(phy97.f.p, Species=="furcata")
pfur97bs.f.p <- subset_samples(phy97bs.f.p, Species=="furcata")
pfur100.f.p <- subset_samples(phy100.f.p, Species=="furcata")
# Plot custom barplots for Porites furcata
par(mfrow=c(3,1), mar=c(4,4,2,0), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(pfur97.f.p, main="97% OTUs")
otubarplot(pfur97bs.f.p, main="97% within-sample OTUs")
otubarplot(pfur100.f.p, main="100% OTUs")
```
```{r pfur.net, cache=T, fig.height=6, fig.width=6}
# Network analysis
pfur.net <- makenet(pfur97bs.f.p, 0)
par(mar=c(0,0,0,0))
set.seed(54538)
plotnet(pfur.net)
```

@@

## *Orbicella annularis*
```{r oann, cache=F, fig.height=12, fig.width=8}
# Create subsetted phyloseq objects for Orbicella annularis
oann97.f.p <- subset_samples(phy97.f.p, Species=="annularis")
oann97bs.f.p <- subset_samples(phy97bs.f.p, Species=="annularis")
oann100.f.p <- subset_samples(phy100.f.p, Species=="annularis")
# Plot custom barplots for Orbicella annularis
par(mfrow=c(3,1), mar=c(4,4,2,0), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(oann97.f.p, main="97% OTUs")
otubarplot(oann97bs.f.p, main="97% within-sample OTUs")
otubarplot(oann100.f.p, main="100% OTUs")
```
```{r oann.net, cache=T, fig.height=6, fig.width=6}
# Network analysis
oann.net <- makenet(oann97bs.f.p, 0)
par(mar=c(0,0,0,0))
set.seed(54538)
plotnet(oann.net)
```

@@

## *Millepora alcicornis*
```{r malc, cache=T, fig.height=12, fig.width=8}
# Create subsetted phyloseq objects for Millepora alcicornis
malc97.f.p <- subset_samples(phy97.f.p, Species=="alcicornis")
malc97bs.f.p <- subset_samples(phy97bs.f.p, Species=="alcicornis")
malc100.f.p <- subset_samples(phy100.f.p, Species=="alcicornis")
# Plot custom barplots for Millepora alcicornis
par(mfrow=c(3,1), mar=c(4,4,2,0), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(malc97.f.p, main="97% OTUs")
otubarplot(malc97bs.f.p, main="97% within-sample OTUs")
otubarplot(malc100.f.p, main="100% OTUs")
```
```{r malc.net, cache=T, fig.height=6, fig.width=6}
# Network analysis
malc.net <- makenet(malc97bs.f.p, 0)
par(mar=c(0,0,0,0))
set.seed(54538)
plotnet(malc.net)
```

@@

## *Siderastrea siderea*
```{r ssid, cache=T, fig.height=12, fig.width=8}
# Create subsetted phyloseq objects for Siderastrea siderea
ssid97.f.p <- subset_samples(phy97.f.p, Species=="siderea")
ssid97bs.f.p <- subset_samples(phy97bs.f.p, Species=="siderea")
ssid100.f.p <- subset_samples(phy100.f.p, Species=="siderea")
# Plot custom barplots for Siderastrea siderea
par(mfrow=c(3,1), mar=c(4,4,2,0), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(ssid97.f.p, main="97% OTUs")
otubarplot(ssid97bs.f.p, main="97% within-sample OTUs")
otubarplot(ssid100.f.p, main="100% OTUs")
```
```{r ssid.net, cache=T, fig.height=6, fig.width=6}
# Network analysis
ssid.net <- makenet(ssid97bs.f.p, 0)
par(mar=c(0,0,0,0))
set.seed(834)
plotnet(ssid.net)
```

@@

## *Favia fragum*
```{r ffra, cache=T, fig.height=12, fig.width=8}
# Create subsetted phyloseq objects for Favia fragum
ffra97.f.p <- subset_samples(phy97.f.p, Species=="fragum")
ffra97bs.f.p <- subset_samples(phy97bs.f.p, Species=="fragum")
ffra100.f.p <- subset_samples(phy100.f.p, Species=="fragum")
# Plot custom barplots for Favia fragum
par(mfrow=c(3,1), mar=c(4,4,2,0), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(ffra97.f.p, main="97% OTUs")
otubarplot(ffra97bs.f.p, main="97% within-sample OTUs")
otubarplot(ffra100.f.p, main="100% OTUs")
```
```{r ffra.net, cache=T, fig.height=6, fig.width=6}
# Network analysis
ffra.net <- makenet(ffra97bs.f.p, 0)
par(mar=c(0,0,0,0))
set.seed(54538)
plotnet(ffra.net)
```

@@

## *Siderastrea radians*
```{r srad, cache=T, fig.height=12, fig.width=8}
# Create subsetted phyloseq objects for Siderastrea radians
srad97.f.p <- subset_samples(phy97.f.p, Species=="radians")
srad97bs.f.p <- subset_samples(phy97bs.f.p, Species=="radians")
srad100.f.p <- subset_samples(phy100.f.p, Species=="radians")
# Plot custom barplots for Siderastrea radians
par(mfrow=c(3,1), mar=c(4,4,2,0), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(srad97.f.p, main="97% OTUs")
otubarplot(srad97bs.f.p, main="97% within-sample OTUs")
otubarplot(srad100.f.p, main="100% OTUs")
```
```{r srad.net, cache=T, fig.height=6, fig.width=6}
# Network analysis
srad.net <- makenet(srad97bs.f.p, 0)
par(mar=c(0,0,0,0))
set.seed(54538)
plotnet(srad.net)
```

>>>>

# Data pre-processing
### Import dataset
The [phyloseq](https://joey711.github.io/phyloseq/) package is used here to combine the OTU table, taxonomic assignments, and sample metadata into a single R data object (class 'phyloseq') to facilitate downstream analyses.
```{r data_import, cache=T}
# Import OTU tables
OTU97 <- otu_table(read.table("data/97_otus.tsv", header=T, check.names=F, row.names=1,
                              skip=1, sep="\t", comment.char=""), taxa_are_rows=T)
OTU97bs <- otu_table(read.table("data/97_otus_bysample.tsv", header=T, check.names=F, row.names=1,
                              skip=1, sep="\t", comment.char=""), taxa_are_rows=T)
OTU97bsp <- otu_table(read.table("data/97_otus_byspecies.tsv", header=T, check.names=F, row.names=1,
                              skip=1, sep="\t", comment.char=""), taxa_are_rows=T)
OTU100 <- otu_table(read.table("data/100_otus.tsv", header=T, check.names=F, row.names=1,
                               skip=1, sep="\t", comment.char=""), taxa_are_rows=T)
# Import taxonomy
TAX97 <- tax_table(as.matrix(read.table("data/97_otus_tax_assignments.txt", sep="\t", row.names=2,
                                        col.names=c("Clade", "OTU.ID", "Taxonomy", "Eval", "Subtype"))))
TAX97bs <- tax_table(as.matrix(read.table("data/97_otus_bysample_tax_assignments.txt", sep="\t", row.names=2,
                                        col.names=c("Clade", "OTU.ID", "Taxonomy", "Eval", "Subtype"))))
TAX97bsp <- tax_table(as.matrix(read.table("data/97_otus_byspecies_tax_assignments.txt", sep="\t", row.names=2,
                                        col.names=c("Clade", "OTU.ID", "Taxonomy", "Eval", "Subtype"))))
TAX100 <- tax_table(as.matrix(read.table("data/100_otus_tax_assignments.txt", sep="\t", row.names=2,
                                         col.names=c("Clade", "OTU.ID", "Taxonomy", "Eval", "Subtype"))))
# Import sample data
SAM <- sample_data(read.delim("data/mapping_file.txt", sep="\t", header=T, row.names=1))
# Build phyloseq objects
phy97 <- phyloseq(OTU97, TAX97, SAM)
phy97bs <- phyloseq(OTU97bs, TAX97bs, SAM)
phy97bsp <- phyloseq(OTU97bsp, TAX97bsp, SAM)
phy100 <- phyloseq(OTU100, TAX100, SAM)
```

### Filter dataset
```{r filter_dataset, cache=T}
# Filter OTUs by minimum count
# Set threshold count
n <- 10
# Identify OTUs below threshold count
taxa97 <- taxa_sums(phy97)[which(taxa_sums(phy97) >= n)]
taxa97bs <- taxa_sums(phy97bs)[which(taxa_sums(phy97bs) >= n)]
taxa97bsp <- taxa_sums(phy97bsp)[which(taxa_sums(phy97bsp) >= n)] 
taxa100 <- taxa_sums(phy100)[which(taxa_sums(phy100) >= n)]
# Remove taxa below threshold count
phy97.f <- prune_taxa(names(taxa97), phy97)
phy97bs.f <- prune_taxa(names(taxa97bs), phy97bs)
phy97bsp.f <- prune_taxa(names(taxa97bsp), phy97bsp)
phy100.f <- prune_taxa(names(taxa100), phy100)

# Filter samples by minimum count
# Set threshold number of reads
sn <- 200
# Remove samples with fewer reads than threshold
phy97.f <- prune_samples(sample_sums(phy97.f)>=sn, phy97.f)
phy97bs.f <- prune_samples(sample_sums(phy97bs.f)>=sn, phy97bs.f)
phy97bsp.f <- prune_samples(sample_sums(phy97bsp.f)>=sn, phy97bsp.f)
phy100.f <- prune_samples(sample_sums(phy100.f)>=sn, phy100.f)
```
* Minimum count to retain OTU: `r n`
* Minimum count to retain sample: `r sn`

@@

#### Filtered dataset
```{r phy.f_histograms, cache=T, fig.height=4, fig.width=8}
# Compute summary statistics
stats97.f <- data.frame(`97% OTUs`=t(phystats(phy97.f)), check.names=F)
stats97bs.f <- data.frame(`97% within-sample OTUs`=t(phystats(phy97bs.f)), check.names=F)
stats97bsp.f <- data.frame(`97% within-species OTUs`=t(phystats(phy97bsp.f)), check.names=F)
stats100.f <- data.frame(`100% OTUs`=t(phystats(phy100.f)), check.names=F)

# Create and plot histograms
taxhist97 <- hist(log10(taxa_sums(phy97.f)), plot=F)
taxhist97bs <- hist(log10(taxa_sums(phy97bs.f)), plot=F)
taxhist97bsp <- hist(log10(taxa_sums(phy97bsp.f)), plot=F)
taxhist100 <- hist(log10(taxa_sums(phy100.f)), plot=F)
samhist97 <- hist(log10(sample_sums(phy97.f)), plot=F)
samhist97bs <- hist(log10(sample_sums(phy97bs.f)), plot=F)
samhist97bsp <- hist(log10(sample_sums(phy97bsp.f)), plot=F)
samhist100 <- hist(log10(sample_sums(phy100.f)), plot=F)

par(mfrow=c(2, 4), mar=c(3,3,1,1))
plot(taxhist97, col="black", main="97% OTU counts", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. OTUs", cex.lab=0.75, cex.axis=0.75)
plot(taxhist97bs, col="black", main="97% sample OTU counts", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. OTUs", cex.lab=0.75, cex.axis=0.75)
plot(taxhist97bsp, col="black", main="97% species OTU counts", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. OTUs", cex.lab=0.75, cex.axis=0.75)
plot(taxhist100, col="black", main="100% OTU counts", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. OTUs", cex.lab=0.75, cex.axis=0.75)
plot(samhist97, col="black", main="97% OTU reads per sample", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. samples", cex.lab=0.75, cex.axis=0.75)
plot(samhist97bs, col="black", main="97% sample OTU reads per sample", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. samples", cex.lab=0.75, cex.axis=0.75)
plot(samhist97bsp, col="black", main="97% species OTU reads per sample", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. samples", cex.lab=0.75, cex.axis=0.75)
plot(samhist100, col="black", main="100% OTU reads per sample", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. samples", cex.lab=0.75, cex.axis=0.75)

# Create stats table
knitr::kable(cbind(stats97.f, stats97bs.f, stats97bsp.f, stats100.f))
```

#### Raw dataset
```{r phy_stats, cache=T, fig.height=4, fig.width=8}
# Compute summary statistics
stats97 <- data.frame(`97% OTUs`=t(phystats(phy97)), check.names=F)
stats97bs <- data.frame(`97% within-sample OTUs`=t(phystats(phy97bs)), check.names=F)
stats97bsp <- data.frame(`97% within-species OTUs`=t(phystats(phy97bsp)), check.names=F)
stats100 <- data.frame(`100% OTUs`=t(phystats(phy100)), check.names=F)

@@

taxhist97bsp <- hist(log10(taxa_sums(phy97bsp)), plot=F)
samhist97bsp <- hist(log10(sample_sums(phy97bsp)), plot=F)

@@

par(mfrow=c(2, 4), mar=c(3,3,1,1))
plot(taxhist97, col="black", main="97% OTU counts", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. OTUs", cex.lab=0.75, cex.axis=0.75)
plot(taxhist97bs, col="black", main="97% within-sample OTU counts", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. OTUs", cex.lab=0.75, cex.axis=0.75)
plot(taxhist97bsp, col="black", main="97% within-species OTU counts", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. OTUs", cex.lab=0.75, cex.axis=0.75)
plot(taxhist100, col="black", main="100% OTU counts", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. OTUs", cex.lab=0.75, cex.axis=0.75)
plot(samhist97, col="black", main="97% OTU reads per sample", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. samples", cex.lab=0.75, cex.axis=0.75)
plot(samhist97bs, col="black", main="97% within-sample OTU reads per sample", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. samples", cex.lab=0.75, cex.axis=0.75)
plot(samhist97bs, col="black", main="97% within-species OTU reads per sample", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. samples", cex.lab=0.75, cex.axis=0.75)
plot(samhist100, col="black", main="100% OTU reads per sample", xlim=c(0, 6), las=1, mgp=c(2,0.5,0),
     xlab="No. sequences (log10)", ylab="No. samples", cex.lab=0.75, cex.axis=0.75)

# Create stats table
knitr::kable(cbind(stats97, stats97bs, stats97bsp, stats100))
```

### Transform count data
Count data are transformed to both relative abundance (proportions) and square-root proportions for downstream statistical analyses.
```{r transform}
# Convert to proportion (relative abundance)
phy97.f.p <- transform_sample_counts(phy97.f, function(x) x/sum(x))
phy97bs.f.p <- transform_sample_counts(phy97bs.f, function(x) x/sum(x))
phy97bsp.f.p <- transform_sample_counts(phy97bsp.f, function(x) x/sum(x))
phy100.f.p <- transform_sample_counts(phy100.f, function(x) x/sum(x))
# Apply transformation function
transform <- function(x) sqrt(x/sum(x))  # Set transformation function
phy97.f.t <- transform_sample_counts(phy97.f, transform)  # Transform data
phy97bs.f.t <- transform_sample_counts(phy97bs.f, transform) 
phy97bsp.f.t <- transform_sample_counts(phy97bsp.f, transform) 
phy100.f.t <- transform_sample_counts(phy100.f, transform)
```

@@

## *Diploria strigosa*  
Notice how 97% OTUs and 97% within-sample OTUs are **NOT THE SAME!** Specifically, notice how in the 97% OTU dataset, all samples except two are dominated by the same OTU, denovo235. In the 97% within-sample OTU dataset, those samples are dominated by three different OTUs -- denovo73, denovo231, and denovo10. Look at the 100% OTU dataset to see the composition of unique sequence variants in these samples. Indeed, these samples are dominated by **different sequence variants**. However, in the 97% OTU approach, they have all been collapsed into the same OTU, while in the 97% within-sample clustering approach, samples with different dominant sequence variants are maintained as separate OTUs, even though these different sequence variants are more than 97% similar to each other. I suggest that the  within-sample clustering approach is a better way of treating *Symbiodinium* ITS2 sequence data.
```{r dstr, cache=T, fig.height=16, fig.width=8}
# Create subsetted phyloseq objects for Diploria strigosa
dstr97.f.p <- subset_samples(phy97.f.p, Species=="strigosa")
dstr97bs.f.p <- subset_samples(phy97bs.f.p, Species=="strigosa")
dstr97bsp.f.p <- subset_samples(phy97bsp.f.p, Species=="strigosa")
dstr100.f.p <- subset_samples(phy100.f.p, Species=="strigosa")
# Plot custom barplots for Diploria strigosa
par(mfrow=c(4,1), mar=c(4,4,2,0), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(dstr97.f.p, main="97% OTUs")
otubarplot(dstr97bsp.f.p, main="97% within-species OTUs")
otubarplot(dstr97bs.f.p, main="97% within-sample OTUs")
otubarplot(dstr100.f.p, main="100% OTUs")
```

@@

## *Porites furcata*
```{r pfur, cache=T, fig.height=16, fig.width=8}
# Create subsetted phyloseq objects for Porites furcata
pfur97.f.p <- subset_samples(phy97.f.p, Species=="furcata")
pfur97bsp.f.p <- subset_samples(phy97bsp.f.p, Species=="furcata")
pfur97bs.f.p <- subset_samples(phy97bs.f.p, Species=="furcata")
pfur100.f.p <- subset_samples(phy100.f.p, Species=="furcata")
# Plot custom barplots for Porites furcata
par(mfrow=c(4,1), mar=c(4,4,2,0), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(pfur97.f.p, main="97% OTUs")
otubarplot(pfur97bsp.f.p, main="97% within-species OTUs")
otubarplot(pfur97bs.f.p, main="97% within-sample OTUs")
otubarplot(pfur100.f.p, main="100% OTUs")
```
```{r pfur.net, cache=T, fig.height=6, fig.width=6}
# Network analysis
pfur.net <- makenet(pfur97bs.f.p, 0)
par(mar=c(0,0,0,0))
set.seed(54538)
plotnet(pfur.net)
```

@@

## *Orbicella annularis*
```{r oann, cache=F, fig.height=16, fig.width=8}
# Create subsetted phyloseq objects for Orbicella annularis
oann97.f.p <- subset_samples(phy97.f.p, Species=="annularis")
oann97bsp.f.p <- subset_samples(phy97bsp.f.p, Species=="annularis")
oann97bs.f.p <- subset_samples(phy97bs.f.p, Species=="annularis")
oann100.f.p <- subset_samples(phy100.f.p, Species=="annularis")
# Plot custom barplots for Orbicella annularis
par(mfrow=c(4,1), mar=c(4,4,2,0), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(oann97.f.p, main="97% OTUs")
otubarplot(oann97bsp.f.p, main="97% within-species OTUs")
otubarplot(oann97bs.f.p, main="97% within-sample OTUs")
otubarplot(oann100.f.p, main="100% OTUs")
```
```{r oann.net, cache=T, fig.height=6, fig.width=6}
# Network analysis
oann.net <- makenet(oann97bs.f.p, 0)
par(mar=c(0,0,0,0))
set.seed(54538)
plotnet(oann.net)
```

@@

## *Millepora alcicornis*
```{r malc, cache=T, fig.height=16, fig.width=8}
# Create subsetted phyloseq objects for Millepora alcicornis
malc97.f.p <- subset_samples(phy97.f.p, Species=="alcicornis")
malc97bsp.f.p <- subset_samples(phy97bsp.f.p, Species=="alcicornis")
malc97bs.f.p <- subset_samples(phy97bs.f.p, Species=="alcicornis")
malc100.f.p <- subset_samples(phy100.f.p, Species=="alcicornis")
# Plot custom barplots for Millepora alcicornis
par(mfrow=c(4,1), mar=c(4,4,2,0), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(malc97.f.p, main="97% OTUs")
otubarplot(malc97bsp.f.p, main="97% within-sample OTUs")
otubarplot(malc97bs.f.p, main="97% within-sample OTUs")
otubarplot(malc100.f.p, main="100% OTUs")
```
```{r malc.net, cache=T, fig.height=6, fig.width=6}
# Network analysis
malc.net <- makenet(malc97bs.f.p, 0)
par(mar=c(0,0,0,0))
set.seed(54538)
plotnet(malc.net)
```

@@

## *Siderastrea siderea*
```{r ssid, cache=T, fig.height=16, fig.width=8}
# Create subsetted phyloseq objects for Siderastrea siderea
ssid97.f.p <- subset_samples(phy97.f.p, Species=="siderea")
ssid97bsp.f.p <- subset_samples(phy97bsp.f.p, Species=="siderea")
ssid97bs.f.p <- subset_samples(phy97bs.f.p, Species=="siderea")
ssid100.f.p <- subset_samples(phy100.f.p, Species=="siderea")
# Plot custom barplots for Siderastrea siderea
par(mfrow=c(4,1), mar=c(4,4,2,0), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(ssid97.f.p, main="97% OTUs")
otubarplot(ssid97bsp.f.p, main="97% within-species OTUs")
otubarplot(ssid97bs.f.p, main="97% within-sample OTUs")
otubarplot(ssid100.f.p, main="100% OTUs")
```
```{r ssid.net, cache=T, fig.height=6, fig.width=6}
# Network analysis
ssid.net <- makenet(ssid97bs.f.p, 0)
par(mar=c(0,0,0,0))
set.seed(834)
plotnet(ssid.net)
```

@@

## *Favia fragum*
```{r ffra, cache=T, fig.height=16, fig.width=8}
# Create subsetted phyloseq objects for Favia fragum
ffra97.f.p <- subset_samples(phy97.f.p, Species=="fragum")
ffra97bsp.f.p <- subset_samples(phy97bsp.f.p, Species=="fragum")
ffra97bs.f.p <- subset_samples(phy97bs.f.p, Species=="fragum")
ffra100.f.p <- subset_samples(phy100.f.p, Species=="fragum")
# Plot custom barplots for Favia fragum
par(mfrow=c(4,1), mar=c(4,4,2,0), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(ffra97.f.p, main="97% OTUs")
otubarplot(ffra97bsp.f.p, main="97% within-species OTUs")
otubarplot(ffra97bs.f.p, main="97% within-sample OTUs")
otubarplot(ffra100.f.p, main="100% OTUs")
```
```{r ffra.net, cache=T, fig.height=6, fig.width=6}
# Network analysis
ffra.net <- makenet(ffra97bs.f.p, 0)
par(mar=c(0,0,0,0))
set.seed(54538)
plotnet(ffra.net)
```

@@

## *Siderastrea radians*
```{r srad, cache=T, fig.height=16, fig.width=8}
# Create subsetted phyloseq objects for Siderastrea radians
srad97.f.p <- subset_samples(phy97.f.p, Species=="radians")
srad97bsp.f.p <- subset_samples(phy97bsp.f.p, Species=="radians")
srad97bs.f.p <- subset_samples(phy97bs.f.p, Species=="radians")
srad100.f.p <- subset_samples(phy100.f.p, Species=="radians")
# Plot custom barplots for Siderastrea radians
par(mfrow=c(4,1), mar=c(4,4,2,0), mgp=c(0,0.5,0), lwd=0.1)
otubarplot(srad97.f.p, main="97% OTUs")
otubarplot(srad97bsp.f.p, main="97% within-species OTUs")
otubarplot(srad97bs.f.p, main="97% within-sample OTUs")
otubarplot(srad100.f.p, main="100% OTUs")
```
```{r srad.net, cache=T, fig.height=6, fig.width=6}
# Network analysis
srad.net <- makenet(srad97bs.f.p, 0)
par(mar=c(0,0,0,0))
set.seed(54538)
plotnet(srad.net)
```
